---
title: Update condition data
author: 
- affiliation: RACE
  description: Research Fish Biologist
  email: sean.rohan@noaa.gov
  name: Sean Rohan
output: word_document
fontsize: 12pt
addr: 
  l1: 7600 Sand Point Way NE
  l2: NMFS RACE Division, Groundfish Assessment Program
  l3: Seattle, WA 98115
---

# Introduction

This file contains code and instructions for retrieving and updating AFSC/RACE/GAP data in the akfishcondition package. The data in the package are used to calculate the summer groundfish morphometric condition indicators that are used in Ecosystem Status Reports (ESRs) and Ecosystem and Socioeconomic Profiles (ESPs) for the eastern Bering Sea continental shelf, northern Bering Sea, Aleutian Islands, and Gulf of Alaska regions. Data and indicators in the akfishcondition package are intended to be updated annually for each stock assessment cycle.

```{r get_connected, message=FALSE, warning=FALSE}
library(akfishcondition)

channel <- akfishcondition::get_connected()
```


# 1. Run queries to retrieve data for the regions

## 1.1 Aleutian Islands
```{r get_ai, message=FALSE, warning=FALSE}
# Aleutian Islands
ai_biomass <- RODBC::sqlQuery(channel, akfishcondition::sql_to_rqry(system.file("./sql/ai_biomass.sql", package = "akfishcondition"))) %>%
  dplyr::select(-HAUL_COUNT, -CATCH_COUNT, -INPFC_STRATUM_AREA)

ai_lw <- RODBC::sqlQuery(channel, akfishcondition::sql_to_rqry(system.file("./sql/ai_length_weight.sql", package = "akfishcondition"))) %>%
  dplyr::rename(INPFC_STRATUM = INPFC_AREA) %>%
  dplyr::mutate(YEAR = floor(CRUISE/100),
                ID = paste(CRUISE, VESSEL, HAUL, SPECIMENID, sep = "_")) %>%
  dplyr::select(ID, YEAR, REGION, INPFC_STRATUM, COMMON_NAME, SPECIES_CODE, SEX, LENGTH, WEIGHT)

ai_nsamp <- table(ai_lw$INPFC_STRATUM, ai_lw$COMMON_NAME)

ai_dat <- ai_lw %>% 
  dplyr::inner_join(ai_biomass)

# Checks (stratum count match should be TRUE)
ai_count_match <- sum(ai_nsamp == table(ai_dat$INPFC_STRATUM, ai_dat$COMMON_NAME)) == length(ai_nsamp)
print(paste0("AI stratum count match: ", ai_count_match))
print(paste0("Year range: ", min(ai_dat$YEAR), "-",  max(ai_dat$YEAR)))

AI_raw <- list(LW = ai_lw,
               BIOMASS = ai_biomass,
              AKFISHCONDITION_VERSION = utils::packageVersion("akfishcondition"),
              LAST_UPDATE = Sys.Date())

# Subset for ESP
pcod_ai_dat <- ai_dat %>% dplyr::filter(SPECIES_CODE == 21720)
```

## 1.2 Gulf of Alaska
```{r get_goa, message=FALSE, warning=FALSE}
# Retrieve GOA biomass for GOA INPFC strata
goa_biomass <- RODBC::sqlQuery(channel, akfishcondition::sql_to_rqry(system.file("./sql/goa_biomass.sql", package = "akfishcondition"))) %>%
  dplyr::select(-HAUL_COUNT, -CATCH_COUNT, -INPFC_STRATUM_AREA)

# Retrieve GOA length-weight, rename INPFFC_AREA to INPFC_STRATUM and create YEAR field.
goa_lw <- RODBC::sqlQuery(channel, akfishcondition::sql_to_rqry(system.file("./sql/goa_length_weight.sql", package = "akfishcondition"))) %>%
  dplyr::rename(INPFC_STRATUM = INPFC_AREA) %>%
  dplyr::mutate(YEAR = floor(CRUISE/100),
                ID = paste(CRUISE, VESSEL, HAUL, SPECIMENID, sep = "_")) %>%
  dplyr::select(ID, YEAR, REGION, INPFC_STRATUM, SPECIMENID, COMMON_NAME, SPECIES_CODE, SEX, LENGTH, WEIGHT) %>%
  dplyr::filter(YEAR != 1985) # No estimates for 1985

goa_nsamp <- table(goa_lw$INPFC_STRATUM, goa_lw$COMMON_NAME)

goa_dat <- goa_lw %>% 
  dplyr::inner_join(goa_biomass)

# Checks (stratum count match should be TRUE)
goa_count_match <- sum(goa_nsamp == table(goa_dat$INPFC_STRATUM, goa_dat$COMMON_NAME)) == length(goa_nsamp)
print(paste0("GOA stratum count match: ", goa_count_match))
print(paste0("Year range: ", min(goa_dat$YEAR), "-",  max(goa_dat$YEAR)))

GOA_raw <- list(LW = goa_lw,
                BIOMASS = goa_biomass,
                AKFISHCONDITION_VERSION = utils::packageVersion("akfishcondition"),
                LAST_UPDATE = Sys.Date())

# Subset for ESP
pcod_goa_dat <- goa_dat %>% dplyr::filter(SPECIES_CODE == 21720)
```

## 1.3 Eastern Bering Sea continental shelf

```{r get_ebs, message=FALSE, warning=FALSE}
# Retrieve EBS biomass for aggregated strata (1=10, 2=20, 3=31-32, 4=41-43, 5=50, 6=61-62)
ebs_biomass <- RODBC::sqlQuery(channel, akfishcondition::sql_to_rqry(system.file("./sql/ebs_biomass.sql", package = "akfishcondition")))

# Retrieve EBS length-weight, update STRATUM field and create YEAR field.
ebs_lw <- RODBC::sqlQuery(channel, akfishcondition::sql_to_rqry(system.file("./sql/ebs_length_weight.sql", package = "akfishcondition"))) %>%
  dplyr::mutate(YEAR = floor(CRUISE/100),
                STRATUM = floor(STRATUM/10),
                ID = paste(CRUISE, VESSEL, HAUL, SPECIMENID, sep = "_")) %>%
  dplyr::select(ID, YEAR, REGION, STRATUM, SPECIMENID, COMMON_NAME, SPECIES_CODE, SEX, LENGTH, WEIGHT)

ebs_nsamp <- table(ebs_lw$STRATUM, ebs_lw$COMMON_NAME)

ebs_dat <- ebs_lw %>% 
  dplyr::inner_join(ebs_biomass)

# Checks (stratum count match should be TRUE)
ebs_count_match <- sum(ebs_nsamp == table(ebs_dat$STRATUM, ebs_dat$COMMON_NAME)) == length(ebs_nsamp)
print(paste0("EBS stratum count match: ", ebs_count_match))
print(paste0("Year range: ", min(ebs_dat$YEAR), "-",  max(ebs_dat$YEAR)))

EBS_raw <- list(LW = ebs_lw,
                BIOMASS = ebs_biomass,
                AKFISHCONDITION_VERSION = utils::packageVersion("akfishcondition"),
                LAST_UPDATE = Sys.Date())

# Subset for ESP
pcod_ebs_dat <- ebs_dat %>% dplyr::filter(SPECIES_CODE == 21720)
```


## 1.4 Northern Bering Sea

The northern Bering Sea does not use biomass because the entire region is treated as a single area for the condition indicator.

```{r get_nbs, message=FALSE, warning=FALSE}
# Retrieve NBS length-weight, update STRATUM field and create YEAR field.
nbs_dat <- RODBC::sqlQuery(channel, akfishcondition::sql_to_rqry(system.file("./sql/nbs_length_weight.sql", package = "akfishcondition"))) %>%
  dplyr::mutate(YEAR = floor(CRUISE/100),
                STRATUM = 999,
                ID = paste(CRUISE, VESSEL, HAUL, SPECIMENID, sep = "_")) %>%
  dplyr::select(ID, YEAR, REGION, STRATUM, SPECIMENID, COMMON_NAME, SPECIES_CODE, SEX, LENGTH, WEIGHT)

print(paste0("Year range: ", min(nbs_dat$YEAR), "-",  max(nbs_dat$YEAR)))

NBS_raw <- list(LW = nbs_dat,
                biomass = NA,
                LAST_UPDATE = Sys.Date())

# Subset for ESP
pcod_nbs_dat <- nbs_dat %>% dplyr::filter(SPECIES_CODE == 21720)
```
# 2. Save raw data to inst/extdata

```{r save_raw, message=FALSE, warning=FALSE}
save(AI_raw,
     GOA_raw,
     EBS_raw,
     NBS_raw,
     file = here::here("inst", "extdata", "raw_lw_bio.rda"))
```

# 3. Update regional indicators for ESR and species-specific indictors for ESP

## 2.1 Aleutian Islands
```{r update_ai, message=FALSE, warning=FALSE}
ai_dat$SPECIES_CODE[ai_dat$SPECIES_CODE == 21740  & ai_dat$LENGTH >= 100 & ai_dat$LENGTH <= 250] <- 21741
ai_dat$COMMON_NAME[ai_dat$SPECIES_CODE == 21740] <- "walleye pollock (>250 mm)"
ai_dat$COMMON_NAME[ai_dat$SPECIES_CODE == 21741] <- "walleye pollock (100–250 mm)"

##AI Calculations and figures
ai_spp_vec <- unique(ai_dat$SPECIES_CODE)

# Calculate length weight residuals
for(i in 1:length(ai_spp_vec)) {
  # Separate slope for each stratum. Bias correction according to Brodziak, no outlier detection.
  ai_df <- akfishcondition::calc_lw_residuals(len = ai_dat$LENGTH[ai_dat$SPECIES_CODE == ai_spp_vec[i]], 
                                               wt = ai_dat$WEIGHT[ai_dat$SPECIES_CODE == ai_spp_vec[i]], 
                                               year = ai_dat$YEAR[ai_dat$SPECIES_CODE == ai_spp_vec[i]],
                                               stratum = ai_dat$INPFC_STRATUM[ai_dat$SPECIES_CODE == ai_spp_vec[i]],
                                               make_diagnostics = TRUE, # Make diagnostics
                                               bias.correction = TRUE, # Bias correction turned on
                                               outlier.rm = FALSE, # Outlier removal turned off
                                               region = "AI",
                                               species_code = ai_dat$SPECIES_CODE[ai_dat$SPECIES_CODE == ai_spp_vec[i]])
  
  ai_dat$resid_mean[ai_dat$SPECIES_CODE == ai_spp_vec[i]] <- ai_df$lw.res_mean
  ai_dat$resid_lwr[ai_dat$SPECIES_CODE == ai_spp_vec[i]] <- ai_df$lw.res_lwr
  ai_dat$resid_upr[ai_dat$SPECIES_CODE == ai_spp_vec[i]] <- ai_df$lw.res_upr
  
}

# Estimate mean and std. err for each stratum, filter out strata with less than 10 samples
ai_stratum_resids <- ai_dat %>% 
  dplyr::group_by(COMMON_NAME, SPECIES_CODE, YEAR, INPFC_STRATUM, AREA_BIOMASS) %>%
  dplyr::summarise(stratum_resid_mean = mean(resid_mean),
                   stratum_resid_sd = sd(resid_mean),
                   n = n()) %>%
  dplyr::filter(n >= 10) %>%
  dplyr::mutate(stratum_resid_se = stratum_resid_sd/sqrt(n))

# Weight strata by biomass
for(i in 1:length(ai_spp_vec)) {
  ai_stratum_resids$weighted_resid_mean[ai_stratum_resids$SPECIES_CODE == ai_spp_vec[i]] <- 
    akfishcondition::weight_lw_residuals(residuals = ai_stratum_resids$stratum_resid_mean[ai_stratum_resids$SPECIES_CODE == ai_spp_vec[i]], 
                                         year = ai_stratum_resids$YEAR[ai_stratum_resids$SPECIES_CODE == ai_spp_vec[i]], 
                                         stratum = ai_stratum_resids$INPFC_STRATUM[ai_stratum_resids$SPECIES_CODE == ai_spp_vec[i]], 
                                         stratum_biomass = ai_stratum_resids$AREA_BIOMASS[ai_stratum_resids$SPECIES_CODE == ai_spp_vec[i]])
  ai_stratum_resids$weighted_resid_se[ai_stratum_resids$SPECIES_CODE == ai_spp_vec[i]] <- 
    akfishcondition::weight_lw_residuals(residuals = ai_stratum_resids$stratum_resid_se[ai_stratum_resids$SPECIES_CODE == ai_spp_vec[i]], 
                                         year = ai_stratum_resids$YEAR[ai_stratum_resids$SPECIES_CODE == ai_spp_vec[i]], 
                                         stratum = ai_stratum_resids$INPFC_STRATUM[ai_stratum_resids$SPECIES_CODE == ai_spp_vec[i]], 
                                         stratum_biomass = ai_stratum_resids$AREA_BIOMASS[ai_stratum_resids$SPECIES_CODE == ai_spp_vec[i]])
}

# Biomass-weighted residual and SE by year
ai_ann_mean_resid_df <- ai_stratum_resids %>% 
  dplyr::group_by(YEAR, COMMON_NAME) %>%
  dplyr::summarise(mean_wt_resid = mean(weighted_resid_mean),
                   se_wt_resid = mean(weighted_resid_se))

ai_stratum_resids <- ai_stratum_resids %>%
  dplyr::select(-AREA_BIOMASS, - stratum_resid_sd)
```

## 2.2 Gulf of Alaska
```{r update_goa, message=FALSE, warning=FALSE}
# Convert 10-25 cm lengths to age-1
goa_dat$SPECIES_CODE[goa_dat$SPECIES_CODE == 21740  & goa_dat$LENGTH >= 100 & goa_dat$LENGTH <= 250] <- 21741
goa_dat$COMMON_NAME[goa_dat$SPECIES_CODE == 21740] <- "walleye pollock (>250 mm)"
goa_dat$COMMON_NAME[goa_dat$SPECIES_CODE == 21741] <- "walleye pollock (100–250 mm)"

## Calculations and figures
goa_spp_vec <- unique(goa_dat$SPECIES_CODE)

# Calculate length weight residuals
for(i in 1:length(goa_spp_vec)) {
  # Separate slope for each stratum. Bias correction according to Brodziak, no outlier detection.
  goa_df <- akfishcondition::calc_lw_residuals(len = goa_dat$LENGTH[goa_dat$SPECIES_CODE == goa_spp_vec[i]], 
                                               wt = goa_dat$WEIGHT[goa_dat$SPECIES_CODE == goa_spp_vec[i]], 
                                               year = goa_dat$YEAR[goa_dat$SPECIES_CODE == goa_spp_vec[i]],
                                               stratum = goa_dat$INPFC_STRATUM[goa_dat$SPECIES_CODE == goa_spp_vec[i]],
                                               make_diagnostics = TRUE, # Make diagnostics
                                               bias.correction = TRUE, # Bias correction turned on
                                               outlier.rm = FALSE, # Outlier removal turned off
                                               region = "GOA",
                                               species_code = goa_dat$SPECIES_CODE[goa_dat$SPECIES_CODE == goa_spp_vec[i]])
  
  goa_dat$resid_mean[goa_dat$SPECIES_CODE == goa_spp_vec[i]] <- goa_df$lw.res_mean
  goa_dat$resid_lwr[goa_dat$SPECIES_CODE == goa_spp_vec[i]] <- goa_df$lw.res_lwr
  goa_dat$resid_upr[goa_dat$SPECIES_CODE == goa_spp_vec[i]] <- goa_df$lw.res_upr
  
}

# Estimate mean and std. err for each stratum, filter out strata with less than 10 samples
goa_stratum_resids <- goa_dat %>% 
  dplyr::group_by(COMMON_NAME, SPECIES_CODE, YEAR, INPFC_STRATUM, AREA_BIOMASS) %>%
  dplyr::summarise(stratum_resid_mean = mean(resid_mean),
                   stratum_resid_sd = sd(resid_mean),
                   n = n()) %>%
  dplyr::filter(n >= 10) %>%
  dplyr::mutate(stratum_resid_se = stratum_resid_sd/sqrt(n))

# Weight strata by biomass
for(i in 1:length(goa_spp_vec)) {
  goa_stratum_resids$weighted_resid_mean[goa_stratum_resids$SPECIES_CODE == goa_spp_vec[i]] <- 
    akfishcondition::weight_lw_residuals(residuals = goa_stratum_resids$stratum_resid_mean[goa_stratum_resids$SPECIES_CODE == goa_spp_vec[i]], 
                                         year = goa_stratum_resids$YEAR[goa_stratum_resids$SPECIES_CODE == goa_spp_vec[i]], 
                                         stratum = goa_stratum_resids$INPFC_STRATUM[goa_stratum_resids$SPECIES_CODE == goa_spp_vec[i]], 
                                         stratum_biomass = goa_stratum_resids$AREA_BIOMASS[goa_stratum_resids$SPECIES_CODE == goa_spp_vec[i]])
  goa_stratum_resids$weighted_resid_se[goa_stratum_resids$SPECIES_CODE == goa_spp_vec[i]] <- 
    akfishcondition::weight_lw_residuals(residuals = goa_stratum_resids$stratum_resid_se[goa_stratum_resids$SPECIES_CODE == goa_spp_vec[i]], 
                                         year = goa_stratum_resids$YEAR[goa_stratum_resids$SPECIES_CODE == goa_spp_vec[i]], 
                                         stratum = goa_stratum_resids$INPFC_STRATUM[goa_stratum_resids$SPECIES_CODE == goa_spp_vec[i]], 
                                         stratum_biomass = goa_stratum_resids$AREA_BIOMASS[goa_stratum_resids$SPECIES_CODE == goa_spp_vec[i]])
}

# Biomass-weighted residual and SE by year
goa_ann_mean_resid_df <- goa_stratum_resids %>% 
  dplyr::group_by(YEAR, COMMON_NAME) %>%
  dplyr::summarise(mean_wt_resid = mean(weighted_resid_mean),
                   se_wt_resid = mean(weighted_resid_se))

goa_stratum_resids <- goa_stratum_resids %>%
  dplyr::select(-AREA_BIOMASS, - stratum_resid_sd)
```

## 2.3 Eastern Bering Sea
```{r update_ebs, message=FALSE, warning=FALSE}
# Convert 10-25 cm lengths to age-1
ebs_dat$SPECIES_CODE[ebs_dat$SPECIES_CODE == 21740  & ebs_dat$LENGTH >= 100 & ebs_dat$LENGTH <= 250] <- 21741
ebs_dat$COMMON_NAME[ebs_dat$SPECIES_CODE == 21741] <- "walleye pollock (100–250 mm)"
ebs_dat$COMMON_NAME[ebs_dat$SPECIES_CODE == 21740] <- "walleye pollock (>250 mm)"

ebs_spp_vec <- unique(ebs_dat$SPECIES_CODE)

# Calculate length weight residuals
for(i in 1:length(ebs_spp_vec)) {
  # Separate slope for each stratum. Bias correction according to Brodziak, no outlier detection.
  ebs_df <- akfishcondition::calc_lw_residuals(len = ebs_dat$LENGTH[ebs_dat$SPECIES_CODE == ebs_spp_vec[i]], 
                                               wt = ebs_dat$WEIGHT[ebs_dat$SPECIES_CODE == ebs_spp_vec[i]], 
                                               year = ebs_dat$YEAR[ebs_dat$SPECIES_CODE == ebs_spp_vec[i]],
                                               stratum = ebs_dat$STRATUM[ebs_dat$SPECIES_CODE == ebs_spp_vec[i]],
                                               make_diagnostics = TRUE, # Make diagnostics
                                               bias.correction = TRUE, # Bias correction turned on
                                               outlier.rm = FALSE, # Outlier removal turned on
                                               region = "EBS",
                                               species_code = ebs_dat$SPECIES_CODE[ebs_dat$SPECIES_CODE == ebs_spp_vec[i]])
  
  ebs_dat$resid_mean[ebs_dat$SPECIES_CODE == ebs_spp_vec[i]] <- ebs_df$lw.res_mean
  ebs_dat$resid_lwr[ebs_dat$SPECIES_CODE == ebs_spp_vec[i]] <- ebs_df$lw.res_lwr
  ebs_dat$resid_upr[ebs_dat$SPECIES_CODE == ebs_spp_vec[i]] <- ebs_df$lw.res_upr
  
}

# Estimate mean and std. err for each stratum, filter out strata with less than 10 samples
ebs_stratum_resids <- ebs_dat %>% 
  dplyr::group_by(COMMON_NAME, SPECIES_CODE, YEAR, STRATUM, BIOMASS) %>%
  dplyr::summarise(stratum_resid_mean = mean(resid_mean),
                   stratum_resid_sd = sd(resid_mean),
                   n = n()) %>%
  dplyr::filter(n >= 10) %>%
  dplyr::mutate(stratum_resid_se = stratum_resid_sd/sqrt(n))

# Weight strata by biomass
for(i in 1:length(ebs_spp_vec)) {
  ebs_stratum_resids$weighted_resid_mean[ebs_stratum_resids$SPECIES_CODE == ebs_spp_vec[i]] <- 
    akfishcondition::weight_lw_residuals(residuals = ebs_stratum_resids$stratum_resid_mean[ebs_stratum_resids$SPECIES_CODE == ebs_spp_vec[i]], 
                                         year = ebs_stratum_resids$YEAR[ebs_stratum_resids$SPECIES_CODE == ebs_spp_vec[i]], 
                                         stratum = ebs_stratum_resids$STRATUM[ebs_stratum_resids$SPECIES_CODE == ebs_spp_vec[i]], 
                                         stratum_biomass = ebs_stratum_resids$BIOMASS[ebs_stratum_resids$SPECIES_CODE == ebs_spp_vec[i]])
  ebs_stratum_resids$weighted_resid_se[ebs_stratum_resids$SPECIES_CODE == ebs_spp_vec[i]] <- 
    akfishcondition::weight_lw_residuals(residuals = ebs_stratum_resids$stratum_resid_se[ebs_stratum_resids$SPECIES_CODE == ebs_spp_vec[i]], 
                                         year = ebs_stratum_resids$YEAR[ebs_stratum_resids$SPECIES_CODE == ebs_spp_vec[i]], 
                                         stratum = ebs_stratum_resids$STRATUM[ebs_stratum_resids$SPECIES_CODE == ebs_spp_vec[i]], 
                                         stratum_biomass = ebs_stratum_resids$BIOMASS[ebs_stratum_resids$SPECIES_CODE == ebs_spp_vec[i]])
}

# Biomass-weighted residual and SE by year
ebs_ann_mean_resid_df <- ebs_stratum_resids %>% 
  dplyr::group_by(YEAR, COMMON_NAME) %>%
  dplyr::summarise(mean_wt_resid = mean(weighted_resid_mean),
                   se_wt_resid = mean(weighted_resid_se))

ebs_stratum_resids <- ebs_stratum_resids %>%
  dplyr::select(-BIOMASS, - stratum_resid_sd)
```

## 2.4 Northern Bering Sea

```{r update_nbs, message=FALSE, warning=FALSE}
nbs_dat$SPECIES_CODE[nbs_dat$SPECIES_CODE == 21740 & nbs_dat$LENGTH >= 100 & nbs_dat$LENGTH <= 250] <- 21741
nbs_dat$COMMON_NAME[nbs_dat$SPECIES_CODE == 21741] <- "walleye pollock (100–250 mm)"
nbs_dat$COMMON_NAME[nbs_dat$SPECIES_CODE == 21740] <- "walleye pollock (>250 mm)"

# Get unique species code combinations
nbs_spp_vec <- unique(nbs_dat$SPECIES_CODE)

# Calculate residuals and weighted residuals
for(i in 1:length(nbs_spp_vec)) {
  
  # Separate slope for each stratum. Bias correction according to Brodziak, no outlier detection.
  nbs_dat$resid[nbs_dat$SPECIES_CODE == nbs_spp_vec[i]] <- 
    akfishcondition::calc_lw_residuals(len = nbs_dat$LENGTH[nbs_dat$SPECIES_CODE == nbs_spp_vec[i]], 
                                       wt = nbs_dat$WEIGHT[nbs_dat$SPECIES_CODE == nbs_spp_vec[i]], 
                                       year = nbs_dat$YEAR[nbs_dat$SPECIES_CODE == nbs_spp_vec[i]],
                                       stratum = NA, # Strata are combined for the NBS
                                       make_diagnostics = FALSE, # Make diagnostics
                                       bias.correction = TRUE, # Bias correction turned on
                                       outlier.rm = FALSE, # Outlier removal turned off
                                       include_ci = FALSE,
                                       region = "NBS",
                                       species_code = nbs_dat$SPECIES_CODE[nbs_dat$SPECIES_CODE == nbs_spp_vec[i]])
}

# Biomass-weighted residuals by year
nbs_ann_mean_resid_df <- nbs_dat %>% 
  dplyr::group_by(COMMON_NAME, YEAR) %>%
  dplyr::summarise(mean_resid = mean(resid, na.rm = TRUE),
                   se = sd(resid, na.rm = TRUE)/n(),
                   n = n()) %>%
  dplyr::filter(n >= 10)
```

## 2.5 ESP Pacific cod

```{r update_esp_cod}
# Split into adult and subadult based on Stark (2007)
pcod_ebs_dat$SPECIES_CODE[pcod_ebs_dat$SPECIES_CODE == 21720  & pcod_ebs_dat$LENGTH < 460] <- 21721
pcod_ebs_dat$COMMON_NAME[pcod_ebs_dat$SPECIES_CODE == 21720] <- "Pacific cod (>460 mm)"
pcod_ebs_dat$COMMON_NAME[pcod_ebs_dat$SPECIES_CODE == 21721] <- "Pacific cod (\u2264460 mm)"

pcod_nbs_dat$SPECIES_CODE[pcod_nbs_dat$SPECIES_CODE == 21720  & pcod_nbs_dat$LENGTH < 460] <- 21721
pcod_nbs_dat$COMMON_NAME[pcod_nbs_dat$SPECIES_CODE == 21720] <- "Pacific cod (>460 mm)"
pcod_nbs_dat$COMMON_NAME[pcod_nbs_dat$SPECIES_CODE == 21721] <- "Pacific cod (\u2264460 mm)"

pcod_ai_dat$SPECIES_CODE[pcod_ai_dat$SPECIES_CODE == 21720  & pcod_ai_dat$LENGTH < 460] <- 21721
pcod_ai_dat$COMMON_NAME[pcod_ai_dat$SPECIES_CODE == 21720] <- "Pacific cod (>460 mm)"
pcod_ai_dat$COMMON_NAME[pcod_ai_dat$SPECIES_CODE == 21721] <- "Pacific cod (\u2264460 mm)"

pcod_goa_dat$SPECIES_CODE[pcod_goa_dat$SPECIES_CODE == 21720 & pcod_goa_dat$LENGTH < 420] <- 21721
pcod_goa_dat$COMMON_NAME[pcod_goa_dat$SPECIES_CODE == 21720] <- "Pacific cod (>420 mm)"
pcod_goa_dat$COMMON_NAME[pcod_goa_dat$SPECIES_CODE == 21721] <- "Pacific cod (\u2264420 mm)"

pcod_ebs_spp_vec <- unique(pcod_ebs_dat$SPECIES_CODE)

# Calculate length weight residuals
for(i in 1:length(pcod_ebs_spp_vec)) {
  # Separate slope for each stratum. Bias correction according to Brodziak, no outlier detection.
  pcod_ebs_df <- akfishcondition::calc_lw_residuals(len = pcod_ebs_dat$LENGTH[pcod_ebs_dat$SPECIES_CODE == pcod_ebs_spp_vec[i]], 
                                               wt = pcod_ebs_dat$WEIGHT[pcod_ebs_dat$SPECIES_CODE == pcod_ebs_spp_vec[i]], 
                                               year = pcod_ebs_dat$YEAR[pcod_ebs_dat$SPECIES_CODE == pcod_ebs_spp_vec[i]],
                                               stratum = pcod_ebs_dat$STRATUM[pcod_ebs_dat$SPECIES_CODE == pcod_ebs_spp_vec[i]],
                                               make_diagnostics = TRUE, # Make diagnostics
                                               bias.correction = TRUE, # Bias correction turned on
                                               outlier.rm = FALSE, # Outlier removal turned on
                                               region = "EBS_PCOD",
                                               species_code = pcod_ebs_dat$SPECIES_CODE[pcod_ebs_dat$SPECIES_CODE == pcod_ebs_spp_vec[i]])
  
  pcod_ebs_dat$resid_mean[pcod_ebs_dat$SPECIES_CODE == pcod_ebs_spp_vec[i]] <- pcod_ebs_df$lw.res_mean
  pcod_ebs_dat$resid_lwr[pcod_ebs_dat$SPECIES_CODE == pcod_ebs_spp_vec[i]] <- pcod_ebs_df$lw.res_lwr
  pcod_ebs_dat$resid_upr[pcod_ebs_dat$SPECIES_CODE == pcod_ebs_spp_vec[i]] <- pcod_ebs_df$lw.res_upr
  
}

# Estimate mean and std. err for each stratum, filter out strata with less than 10 samples
pcod_ebs_stratum_resids <- pcod_ebs_dat %>% 
  dplyr::group_by(COMMON_NAME, SPECIES_CODE, YEAR, STRATUM, BIOMASS) %>%
  dplyr::summarise(stratum_resid_mean = mean(resid_mean),
                   stratum_resid_sd = sd(resid_mean),
                   n = n()) %>%
  dplyr::filter(n >= 10) %>%
  dplyr::mutate(stratum_resid_se = stratum_resid_sd/sqrt(n))

# Weight strata by biomass
for(i in 1:length(pcod_ebs_spp_vec)) {
  pcod_ebs_stratum_resids$weighted_resid_mean[pcod_ebs_stratum_resids$SPECIES_CODE == pcod_ebs_spp_vec[i]] <- 
    akfishcondition::weight_lw_residuals(residuals = pcod_ebs_stratum_resids$stratum_resid_mean[pcod_ebs_stratum_resids$SPECIES_CODE == pcod_ebs_spp_vec[i]], 
                                         year = pcod_ebs_stratum_resids$YEAR[pcod_ebs_stratum_resids$SPECIES_CODE == pcod_ebs_spp_vec[i]], 
                                         stratum = pcod_ebs_stratum_resids$STRATUM[pcod_ebs_stratum_resids$SPECIES_CODE == pcod_ebs_spp_vec[i]], 
                                         stratum_biomass = pcod_ebs_stratum_resids$BIOMASS[pcod_ebs_stratum_resids$SPECIES_CODE == pcod_ebs_spp_vec[i]])
  pcod_ebs_stratum_resids$weighted_resid_se[pcod_ebs_stratum_resids$SPECIES_CODE == pcod_ebs_spp_vec[i]] <- 
    akfishcondition::weight_lw_residuals(residuals = pcod_ebs_stratum_resids$stratum_resid_se[pcod_ebs_stratum_resids$SPECIES_CODE == pcod_ebs_spp_vec[i]], 
                                         year = pcod_ebs_stratum_resids$YEAR[pcod_ebs_stratum_resids$SPECIES_CODE == pcod_ebs_spp_vec[i]], 
                                         stratum = pcod_ebs_stratum_resids$STRATUM[pcod_ebs_stratum_resids$SPECIES_CODE == pcod_ebs_spp_vec[i]], 
                                         stratum_biomass = pcod_ebs_stratum_resids$BIOMASS[pcod_ebs_stratum_resids$SPECIES_CODE == pcod_ebs_spp_vec[i]])
}

# Biomass-weighted residual and SE by year
pcod_ebs_ann_mean_resid_df <- pcod_ebs_stratum_resids %>% 
  dplyr::group_by(YEAR, COMMON_NAME) %>%
  dplyr::summarise(mean_wt_resid = mean(weighted_resid_mean),
                   se_wt_resid = mean(weighted_resid_se))

#--------------------------------------
# Northern Bering Sea
#--------------------------------------
# Get unique species code combinations
pcod_nbs_spp_vec <- unique(pcod_nbs_dat$SPECIES_CODE)

# Calculate residuals and weighted residuals
for(i in 1:length(pcod_nbs_spp_vec)) {
  
  # Separate slope for each stratum. Bias correction according to Brodziak, no outlier detection.
  pcod_nbs_dat$resid[pcod_nbs_dat$SPECIES_CODE == pcod_nbs_spp_vec[i]] <- 
    akfishcondition::calc_lw_residuals(len = pcod_nbs_dat$LENGTH[pcod_nbs_dat$SPECIES_CODE == pcod_nbs_spp_vec[i]], 
                                       wt = pcod_nbs_dat$WEIGHT[pcod_nbs_dat$SPECIES_CODE == pcod_nbs_spp_vec[i]], 
                                       year = pcod_nbs_dat$YEAR[pcod_nbs_dat$SPECIES_CODE == pcod_nbs_spp_vec[i]],
                                       stratum = NA, # Strata are combined for the NBS
                                       make_diagnostics = FALSE, # Make diagnostics
                                       bias.correction = TRUE, # Bias correction turned on
                                       outlier.rm = FALSE, # Outlier removal turned off
                                       include_ci = FALSE,
                                       region = "NBS_PCOD",
                                       species_code = pcod_nbs_dat$SPECIES_CODE[pcod_nbs_dat$SPECIES_CODE == pcod_nbs_spp_vec[i]])
}

# Biomass-weighted residuals by year
pcod_nbs_ann_mean_resid_df <- pcod_nbs_dat %>% 
  dplyr::group_by(COMMON_NAME, YEAR, SPECIES_CODE) %>%
  dplyr::summarise(mean_resid = mean(resid, na.rm = TRUE),
                   se = sd(resid, na.rm = TRUE)/n(),
                   n = n()) %>%
  dplyr::filter(n >=10)

#--------------------------------------
# Gulf of Alaska
#--------------------------------------

pcod_goa_spp_vec <- unique(pcod_goa_dat$SPECIES_CODE)

# Calculate length weight residuals
for(i in 1:length(pcod_goa_spp_vec)) {
  # Separate slope for each stratum. Bias correction according to Brodziak, no outlier detection.
  pcod_goa_df <- akfishcondition::calc_lw_residuals(len = pcod_goa_dat$LENGTH[pcod_goa_dat$SPECIES_CODE == pcod_goa_spp_vec[i]], 
                                               wt = pcod_goa_dat$WEIGHT[pcod_goa_dat$SPECIES_CODE == pcod_goa_spp_vec[i]], 
                                               year = pcod_goa_dat$YEAR[pcod_goa_dat$SPECIES_CODE == pcod_goa_spp_vec[i]],
                                               stratum = pcod_goa_dat$INPFC_STRATUM[pcod_goa_dat$SPECIES_CODE == pcod_goa_spp_vec[i]],
                                               make_diagnostics = TRUE, # Make diagnostics
                                               bias.correction = TRUE, # Bias correction turned on
                                               outlier.rm = FALSE, # Outlier removal turned off
                                               region = "GOA_PCOD",
                                               species_code = pcod_goa_dat$SPECIES_CODE[pcod_goa_dat$SPECIES_CODE == pcod_goa_spp_vec[i]])
  
  pcod_goa_dat$resid_mean[pcod_goa_dat$SPECIES_CODE == pcod_goa_spp_vec[i]] <- pcod_goa_df$lw.res_mean
  pcod_goa_dat$resid_lwr[pcod_goa_dat$SPECIES_CODE == pcod_goa_spp_vec[i]] <- pcod_goa_df$lw.res_lwr
  pcod_goa_dat$resid_upr[pcod_goa_dat$SPECIES_CODE == pcod_goa_spp_vec[i]] <- pcod_goa_df$lw.res_upr
  
}

# Estimate mean and std. err for each stratum, filter out strata with less than 10 samples
pcod_goa_stratum_resids <- pcod_goa_dat %>% 
  dplyr::group_by(COMMON_NAME, SPECIES_CODE, YEAR, INPFC_STRATUM, AREA_BIOMASS) %>%
  dplyr::summarise(stratum_resid_mean = mean(resid_mean),
                   stratum_resid_sd = sd(resid_mean),
                   n = n()) %>%
  dplyr::filter(n >= 10) %>%
  dplyr::mutate(stratum_resid_se = stratum_resid_sd/sqrt(n))

# Weight strata by biomass
for(i in 1:length(pcod_goa_spp_vec)) {
  pcod_goa_stratum_resids$weighted_resid_mean[pcod_goa_stratum_resids$SPECIES_CODE == pcod_goa_spp_vec[i]] <- 
    akfishcondition::weight_lw_residuals(residuals = pcod_goa_stratum_resids$stratum_resid_mean[pcod_goa_stratum_resids$SPECIES_CODE == pcod_goa_spp_vec[i]], 
                                         year = pcod_goa_stratum_resids$YEAR[pcod_goa_stratum_resids$SPECIES_CODE == pcod_goa_spp_vec[i]], 
                                         stratum = pcod_goa_stratum_resids$INPFC_STRATUM[pcod_goa_stratum_resids$SPECIES_CODE == pcod_goa_spp_vec[i]], 
                                         stratum_biomass = pcod_goa_stratum_resids$AREA_BIOMASS[pcod_goa_stratum_resids$SPECIES_CODE == pcod_goa_spp_vec[i]])
  pcod_goa_stratum_resids$weighted_resid_se[pcod_goa_stratum_resids$SPECIES_CODE == pcod_goa_spp_vec[i]] <- 
    akfishcondition::weight_lw_residuals(residuals = pcod_goa_stratum_resids$stratum_resid_se[pcod_goa_stratum_resids$SPECIES_CODE == pcod_goa_spp_vec[i]], 
                                         year = pcod_goa_stratum_resids$YEAR[pcod_goa_stratum_resids$SPECIES_CODE == pcod_goa_spp_vec[i]], 
                                         stratum = pcod_goa_stratum_resids$INPFC_STRATUM[pcod_goa_stratum_resids$SPECIES_CODE == pcod_goa_spp_vec[i]], 
                                         stratum_biomass = pcod_goa_stratum_resids$AREA_BIOMASS[pcod_goa_stratum_resids$SPECIES_CODE == pcod_goa_spp_vec[i]])
}

# Biomass-weighted residual and SE by year
pcod_goa_ann_mean_resid_df <- pcod_goa_stratum_resids %>% 
  dplyr::group_by(YEAR, COMMON_NAME) %>%
  dplyr::summarise(mean_wt_resid = mean(weighted_resid_mean),
                   se_wt_resid = mean(weighted_resid_se))

#--------------------------------------
# Aleutian Islands
#--------------------------------------

pcod_ai_spp_vec <- unique(pcod_ai_dat$SPECIES_CODE)

# Calculate length weight residuals
for(i in 1:length(pcod_ai_spp_vec)) {
  # Separate slope for each stratum. Bias correction according to Brodziak, no outlier detection.
  pcod_ai_df <- akfishcondition::calc_lw_residuals(len = pcod_ai_dat$LENGTH[pcod_ai_dat$SPECIES_CODE == pcod_ai_spp_vec[i]], 
                                               wt = pcod_ai_dat$WEIGHT[pcod_ai_dat$SPECIES_CODE == pcod_ai_spp_vec[i]], 
                                               year = pcod_ai_dat$YEAR[pcod_ai_dat$SPECIES_CODE == pcod_ai_spp_vec[i]],
                                               stratum = pcod_ai_dat$INPFC_STRATUM[pcod_ai_dat$SPECIES_CODE == pcod_ai_spp_vec[i]],
                                               make_diagnostics = TRUE, # Make diagnostics
                                               bias.correction = TRUE, # Bias correction turned on
                                               outlier.rm = FALSE, # Outlier removal turned off
                                               region = "AI_PCOD",
                                               species_code = pcod_ai_dat$SPECIES_CODE[pcod_ai_dat$SPECIES_CODE == pcod_ai_spp_vec[i]])
  
  pcod_ai_dat$resid_mean[pcod_ai_dat$SPECIES_CODE == pcod_ai_spp_vec[i]] <- pcod_ai_df$lw.res_mean
  pcod_ai_dat$resid_lwr[pcod_ai_dat$SPECIES_CODE == pcod_ai_spp_vec[i]] <- pcod_ai_df$lw.res_lwr
  pcod_ai_dat$resid_upr[pcod_ai_dat$SPECIES_CODE == pcod_ai_spp_vec[i]] <- pcod_ai_df$lw.res_upr
  
}

# Estimate mean and std. err for each stratum, filter out strata with less than 10 samples
pcod_ai_stratum_resids <- pcod_ai_dat %>% 
  dplyr::group_by(COMMON_NAME, SPECIES_CODE, YEAR, INPFC_STRATUM, AREA_BIOMASS) %>%
  dplyr::summarise(stratum_resid_mean = mean(resid_mean),
                   stratum_resid_sd = sd(resid_mean),
                   n = n()) %>%
  dplyr::filter(n >= 10) %>%
  dplyr::mutate(stratum_resid_se = stratum_resid_sd/sqrt(n))

# Weight strata by biomass
for(i in 1:length(pcod_ai_spp_vec)) {
  pcod_ai_stratum_resids$weighted_resid_mean[pcod_ai_stratum_resids$SPECIES_CODE == pcod_ai_spp_vec[i]] <- 
    akfishcondition::weight_lw_residuals(residuals = pcod_ai_stratum_resids$stratum_resid_mean[pcod_ai_stratum_resids$SPECIES_CODE == pcod_ai_spp_vec[i]], 
                                         year = pcod_ai_stratum_resids$YEAR[pcod_ai_stratum_resids$SPECIES_CODE == pcod_ai_spp_vec[i]], 
                                         stratum = pcod_ai_stratum_resids$INPFC_STRATUM[pcod_ai_stratum_resids$SPECIES_CODE == pcod_ai_spp_vec[i]], 
                                         stratum_biomass = pcod_ai_stratum_resids$AREA_BIOMASS[pcod_ai_stratum_resids$SPECIES_CODE == pcod_ai_spp_vec[i]])
  pcod_ai_stratum_resids$weighted_resid_se[pcod_ai_stratum_resids$SPECIES_CODE == pcod_ai_spp_vec[i]] <- 
    akfishcondition::weight_lw_residuals(residuals = pcod_ai_stratum_resids$stratum_resid_se[pcod_ai_stratum_resids$SPECIES_CODE == pcod_ai_spp_vec[i]], 
                                         year = pcod_ai_stratum_resids$YEAR[pcod_ai_stratum_resids$SPECIES_CODE == pcod_ai_spp_vec[i]], 
                                         stratum = pcod_ai_stratum_resids$INPFC_STRATUM[pcod_ai_stratum_resids$SPECIES_CODE == pcod_ai_spp_vec[i]], 
                                         stratum_biomass = pcod_ai_stratum_resids$AREA_BIOMASS[pcod_ai_stratum_resids$SPECIES_CODE == pcod_ai_spp_vec[i]])
}

# Biomass-weighted residual and SE by year
pcod_ai_ann_mean_resid_df <- pcod_ai_stratum_resids %>% 
  dplyr::group_by(YEAR, COMMON_NAME) %>%
  dplyr::summarise(mean_wt_resid = mean(weighted_resid_mean),
                   se_wt_resid = mean(weighted_resid_se))
```

# 3. Update version in DESCRIPTION file

Update sysdata.rda with raw data for the condition indicator and write raw data to inst/extdata.

```{r save_to_sysdata, message=FALSE, warning=FALSE}
EBS_INDICATOR <- list(FULL_REGION = as.data.frame(ebs_ann_mean_resid_df),
                      STRATUM = as.data.frame(ebs_stratum_resids),
                      LAST_UPDATE = Sys.Date())
NBS_INDICATOR <- list(FULL_REGION = as.data.frame(nbs_ann_mean_resid_df),
                      STRATUM = NA,
                      LAST_UPDATE = Sys.Date())
GOA_INDICATOR <- list(FULL_REGION = as.data.frame(goa_ann_mean_resid_df),
                      STRATUM = as.data.frame(goa_stratum_resids),
                      LAST_UPDATE = Sys.Date())
AI_INDICATOR <- list(FULL_REGION = as.data.frame(ai_ann_mean_resid_df),
                     STRATUM = as.data.frame(ai_stratum_resids),
                     LAST_UPDATE = Sys.Date())
PCOD_ESP <- list(FULL_REGION_EBS = as.data.frame(pcod_ebs_ann_mean_resid_df),
                 FULL_REGION_GOA = as.data.frame(pcod_goa_ann_mean_resid_df),
                 FULL_REGION_AI = as.data.frame(pcod_ai_ann_mean_resid_df),
                 FULL_REGION_NBS = as.data.frame(pcod_nbs_ann_mean_resid_df),
                 STRATUM_EBS = as.data.frame(pcod_ebs_stratum_resids),
                 STRATUM_GOA = as.data.frame(pcod_goa_stratum_resids),
                 STRATUM_AI = as.data.frame(pcod_ai_stratum_resids),
                 STRATUM_NBS = NA,
                 LAST_UPDATE = Sys.Date())

save(EBS_INDICATOR, 
     NBS_INDICATOR, 
     GOA_INDICATOR,
     AI_INDICATOR,
     PCOD_ESP,
     file = here::here("R", "sysdata.rda"))
```

# 4. Update documentation, install, and restart

Update version number in DESCRIPTION and update documentation using:
```{r updater, eval=FALSE, include=FALSE}
devtools::document(roclets = c('rd', 'collate', 'namespace'))
```

After updating, reinstall the coldpool package:
```{r install, eval=FALSE, include=FALSE}
Rcmd.exe INSTALL --no-multiarch --with-keep.source cold_pool
```

# 5. Check that data appears correctly
Check that the updated cold pool index data is included in the package.
```{r qaqc, eval = TRUE, include =FALSE}
print(akgfmaps::AI_dat$LAST_UPDATE)
print(akgfmaps::GOA_dat$LAST_UPDATE)
print(akgfmaps::EBS_dat$LAST_UPDATE)
print(akgfmaps::NBS_dat$LAST_UPDATE)
```